---
title: "Day2 - Exercise 2"
subtitle: "Example Glucose"
author: "YOUR NAME"
date: "2023-11-10"
output: 
   html_document:
     toc: TRUE
     toc_float: TRUE
---

```{r setup, include = FALSE}
# settings for output
  knitr::opts_chunk$set(echo = TRUE)

# libraries used in this Rmd
  library(ggplot2)
  library(tidyverse)
  library(flextable)
  library(gtsummary)
```

```{r}
# Load Rdata file

 load(file = paste0("../Rdata/", "messy_glucose_prepared_v20231109.Rdata"))

# show general information about dataset

 info$general %>% 
   flextable::flextable()

```



# Data set - Glucose


Glucose tolerance was tested by administering 100g glucose drink

* Glucose was tested before and 1 hour after administering

The data set contains following parameters

* pat_id 
* ...

for ?????????????? patients.

```{r}
# TO DO: Show first few lines of the data set
 


```

# Tables
## Descriptive statistics
* Generate a table with descriptive statistics for the parameters: race, sex, glucose_diff, and age
   - look at homepage of gtsummary: https://www.danieldsjoberg.com/gtsummary/
   - tbl_summary()

```{r}
# descriptive statistics for the parameters: race, sex, glucose_diff, and age
 
```


* Can you generate this table also separated by sex, i.e., descriptive statistics for the parameters race, glucose_diff, and age with one column for man and another for women?
   - Hint: argument 'by'
   

```{r, eval = TRUE, echo = TRUE}
# descriptive statistics by sex

```

* Search online for information how to change the statistics from median (IQR) to mean with standard deviation


## Summary tables
* Sometimes it is interesting to get summary information from your data, e.g., for each category of race you want to know
   - how many man/women
   - what is the minimum age and maximum age observed
   
```{r, eval = TRUE, echo = TRUE}
# extend the code underneath
# use round() to have only 1 decimal place for the columns with age
# use flextable() to get a nicer output

dt_glucose %>%
  group_by(Race) %>% 
  summarize(n_total = n(),
            n_male = sum(Sex == "Male"),
            age_max = max(age)) %>% 
  flextable::flextable()

```
 
### group_by() and mutate()
* Sometimes we need to calculate/create parameters differently for patients, e.g. we want to state if glucose_diff for a patient is above/below average, however, we need to first calculate the mean glucose_diff separately by sex.
   - Hint: look at the examples in documentation of group_by()
   - select the parameters pat_id, Sex, and glucose_diff
   - show data set only for first 6 patients
   - after group_by() it is (often) necessary to 'ungroup' data again

```{r}
dt_glucose %>% 
  select(pat_id, Sex, glucose_diff) 
 
```
  
  

# Plots
**TO DO**
* generate plots which help to look at the data, e.g.,
   - bar charts
   - scatter plot
   - histogram
   - box plot
   - answering questions like: similar age distribution in man and woman?
* save one plot as a file in folder R/Routput using ggsave()   
   


