---
title: "Day2 - Exercise 1"
subtitle: "Save ggplots"
author: "Sereina Herzog"
date: "2023-11-10"
output: 
   html_document:
     toc: TRUE
     toc_float: TRUE
---

```{r setup, include=FALSE}
# settings for output
  knitr::opts_chunk$set(echo = TRUE)

# libraries used in this Rmd
  library(ggplot2)
  library(tidyverse)
  library(flextable)
```


# Iris
## Data set
The data set 'iris' is a famous data set and gives the measurements in centimeters of the variables

* sepal length 
* sepal width
* petal length 
* petal width

for 50 flowers from each of 3 species of iris (*Iris setosa*, *versicolor*, and *virginica*).

```{r dataset1}
# A data set named 'iris'
  dt_iris <- iris 

  dt_iris %>% 
    filter(row_number() < 6) %>% 
    flextable::flextable() 
```

* save datat set 'iris' in 'dt_iris' and show first few lines
   - Hint: Look at Rmd file of Day1 - Exercise 4
* Click on 'dt_iris' in the environment to view the data set within RStudio.


## Plot Iris

```{r}

lab_species <- c("Iris setosa", "Iris virginica", "Iris versicolor")

plot_iris_f <-
  ggplot(data = dt_iris, aes(x = Sepal.Length, y = Sepal.Width, fill = Species)) +
      geom_point(aes(shape = Species), size = 3, alpha = 0.7) +
      
      scale_fill_manual(values = c("darkred", "blue", "orange"),
                        labels = lab_species) +
      scale_shape_manual(values = c(21, 22, 24),
                         labels = lab_species) +
      labs(title = "Dataset Iris") +
      xlab("Sepal length [cm]") +
      ylab("Sepal width [cm]") +
      coord_cartesian(xlim = c(4, 8), ylim = c(2, 5)) +
      theme_bw() +
      theme(plot.title = element_text(face = "bold", size = 25),
            axis.title = element_text(size = 20),
            axis.text = element_text(size = 14),
            legend.title = element_text(size = 16),
            legend.text = element_text(size = 14)) 

plot_iris_f
```


## Save plot
* Save the ggplot using function ggsave() as 'png'
   - use argument 'path' in order to save plot in folder 'R/Routput'
      + Hint: the Rmd file is in folder 'Rmarkdown', i.e. you have to go one level up and then to 'Routput': ../Routput


```{r, eval = TRUE, echo = TRUE}
ggsave(filename = "plot_iris_ggsave_v20231109.png", 
       path = "../Routput/",
       plot = plot_iris_f, 
       units = "cm", width = 12, height = 7, scale = 2)
```

* Try to change/include following arguments in the function and look at the output in folder 'Routput'
   - units
   - scale
   
  



