---
title: "Day1 - Exercise 4"
subtitle: "Getting started with ggplot"
author: "Sereina Herzog"
date: "2023-11-08"
output: 
   html_document:
     toc: TRUE
     toc_float: TRUE
---

```{r setup, include=FALSE}
# settings for output
  knitr::opts_chunk$set(echo = TRUE)

# libraries used in this Rmd
  library(ggplot2)
  library(tidyverse)
  library(flextable)
```

# Changes in Rmarkdown file
* change author name
* add table of content
   - if you use HTML output: add the possibility to have a floating ToC

# Data sets
## Fuel economy data 'mpg'
The data set 'mpg' comes with the *ggplot2* package and contains fuel economy data from 1999 to 2008 for 38 popular models of cars. A data frame with 234 rows and 11 variables. Be aware that 'mpg' has a *tibble* structure (subclass of data.frame). 

```{r dataset1}
# A data set within ggplot named 'mpg'
  dt_mpg <- mpg 

  dt_mpg %>% 
    filter(row_number() < 6) %>% 
    flextable::flextable() 
```

* Use '?mpg' to retrieve information about the variables within the data set.
* Click on 'dt_mpg' in the environment to view the data set within RStudio.

## Pharmacokinetics of Theophylline 'Theoph'
The Theoph data frame has 132 rows and 5 columns of data from an experiment on the pharmacokinetics of theophylline. There are 12 subjects with repeated measurement of theophylline concentration (11x) over time.

```{r dataset2}
# Another data set named 'Theoph'
  dt_theop <- 
     Theoph  %>% 
      as_tibble()

  dt_theop %>% 
    filter(row_number() < 6) %>% 
    flextable::flextable() 
```

* Use '?Theoph' to retrieve information about the variables within the data set.

# Different types of plots using 'dt_mpg'
## Barcharts
Adapt the bar chart illustrating the frequency of manufacturers:

* change color and fill of bars
   - Hint: within 'geom_bar' try something like: fill = "darkgreen"
   
* change background of graph using 'theme_bw()'
   - Hint: look at documentation of theme_bw() - are there other interesting themes?

* we need to change the angle of the labels on the x-axis
   - Hint: search in a search engine with following key words:  ggplot change angle label x axis
   - try different values for angle (e.g., 45)
   - look at documentation of element_text()
      + maybe we need to work with options hjust (and vjust?)
   


```{r}
# SOLUTION
# bar chart
  ggplot(data = dt_mpg, aes(x = manufacturer)) +
    geom_bar(fill = "darkgreen", color = "black") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

* Use the same plot and change from absolute frequencies to relative frequencies.
   - Hint: write 'aes(y = after_stat(prop), group = 1)' in geom_bar()

```{r}
# SOLUTION
# bar chart - copy your solution from above
   ggplot(data = dt_mpg, aes(x = manufacturer)) +
    geom_bar(aes(y = after_stat(prop), group = 1),
                 fill = "darkgreen", color = "black") +
  
    scale_y_continuous(labels = scales::percent) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          axis.title.y = element_blank()) 
 
```

* Do you find a way to change the y axis to display %?
   - Hint: look at documentation of scale_y_continuous()
* Can you get rid of the 'prop' as y axis title?


### Bar chart with figure captions
* Copy solution from above
* For figure captions we have to work with the R chunk options:
   - {r, fig.cap = "YOUR CAPTION"}

```{r, fig.cap = "Figure: Frequency of manufacturers"}
# SOLUTION
# bar chart - copy your solution from above
   ggplot(data = dt_mpg, aes(x = manufacturer)) +
    geom_bar(aes(y = after_stat(prop), group = 1),
                 fill = "darkgreen", color = "black") +
  
    scale_y_continuous(labels = scales::percent) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          axis.title.y = element_blank()) 
 
```


## Histogram
Adapt the histogram illustrating the engine displacement in litres (displ):

* change color and fill of bars
* change background of graph using 'theme_bw()' (or one you prefer)
* we might want to change the interval width into 0.25 liter
   - Hint: look for 'binwidth' in documentation of geom_histogram()
* change x axis title to 'Engine displacement [l]' using xlab()
   - Hint: search in a search engine with following key words:  ggplot change angle label x axis



```{r}
# SOLUTION
# histogram

p_hist_displ <-
  ggplot(data = dt_mpg, aes(x = displ)) +
    geom_histogram(binwidth = 0.25, fill = "darkgreen", color = "black") +
    xlab("Engine displacement [l]") +
    theme_bw() 

p_hist_displ
```

**Note:** The 'p_hist_displ <-' saves the ggplot on the variable name 'p_hist_displ' and therefore we can refer later back to this plot.

### Histogram for each category
* Reuse the histogram code from above (p_hist_displ) and retrieve histograms by "type" of car (class)
   - Hint: facet_grid()

```{r}
# SOLUTION
# histogram by categorical variable

p_hist_displ +
  facet_grid(class ~.)


```


## Boxplot
* Adapt the boxplot underneath such that it illustrates the engine displacement in litres (displ) by  "type" of car (class)
   - Hint: we want to have the "type" of car (class) on the x axis, i.e. aes() needs to be adapted

```{r}
# SOLUTION
# boxplot for displ

p_boxplot_displ <-
 ggplot(data = dt_mpg, aes(y = displ, x = class)) +
    geom_boxplot(fill = "darkgreen", color = "black") +
    ylab("Engine displacement [l]") +
    theme_bw() 

p_boxplot_displ

```



## Combine different plots in one big plot
* We need to install the package 'cowplot' in order to use the function plot_grid()
   - Tools $\rightarrow$ Install packages ...
   - type 'cowplot'
   - press 'Install'
* Adapt the code underneath to 
   - add panel labels
   - have the panels underneath each other


```{r}
# SOLUTION
# Combine different plots in one big plot

 # cowplot::plot_grid(p_boxplot_displ, p_hist_displ, labels = c("A", "B"))
  cowplot::plot_grid(p_boxplot_displ, p_hist_displ, 
                     labels = c("A", "B"), nrow = 2)


```

**Note:** The use of package name together with '::' allows to 

* use a function from a package without loading the package;
* to define which function we want to call in case two/several packages defined a function with the exact same name


## Scatterplot
* Create a scatterplot with engine displacement (displ) on x axis and highway miles per gallon (hwy) on why axis.
   - Hint: use geom_point()
* Adapt axis titles accordingly
* As highway miles per gallon (hwy) is only measured in integers it is likely that we have overlapping observations
   - Hint: use geom_point(alpha = 0.3) and look into documentation for the meaning of 'alpha'
* We want to differentiate if number of cylinders (cyl) is >5 or not
   - Hint: add 'color = (cyl > 5)' to the aes() statement
* Adapt the legend title to 'Number of cylinders'
   - Hints: use labs(color = "ABC") 
* Adapt legend from TRUE/FALSE to >5 and <=5 and the colors used to 'blue' and 'red'
   - Hint: scale_color_manual() and within that labels and values


```{r}
# SOLUTION
# scatterplot

  ggplot(data = dt_mpg, aes(x = displ, y = hwy, color = (cyl > 5))) +
    geom_point(alpha = 0.3) +
  
      xlab("Engine displacement [l]") +
      ylab("Highway miles per gallon") +
      labs(color = "Number of cylinders") +
      scale_color_manual(labels = c(">5", "<=5"),
                         values = c("blue", "red"))+
      theme_bw() 

```

# Different types of plots using 'dt_theop'
## Line plot
* Create a plot with points (geom_point()) and lines (geom_line()) to show theophylline concentration in the subjects over time.
   - Hint: you need to state in aes() that the data is grouped by the patients, i.e. 'group = Subject'
* Adapt axis titles accordingly
* We want to differentiate if dose of theophylline administered orally to the subject (Dose) is >4.5 or not
   - Hint: add 'color = (cyl > 5)' to the aes() statement
   - use your own colors
* Adapt the legend title and legend accordingly
   
```{r}
# SOLUTION
# line plot

  ggplot(data = dt_theop, aes(x = Time, y = conc,
                              group = Subject, color = (Dose > 4.5))) +
    geom_point() +
    geom_line() +
  
      xlab("Time since drug administration  [hr]") +
      ylab("Theophylline concentration in the sample [mg/L]") +
      labs(color = "Dose") +
      scale_color_manual(labels = c(">4.5", "<=4.5"),
                         values = c("purple", "orange"))+
      theme_bw() 

```